```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`rsample` contains a set of functions that can create different types of resamples and corresponding classes for their analysis. 
The goal is to have a modular set of methods that can be used across different R packages for:
 
 * traditional resampling techniques for estimating the sampling distribution of a statistic and
 * estimating model performance using a holdout set
 
The scope of `rsample` is to provide the basic building blocks for creating and analyzing resamples of a data set but does not include code for modeling or calculating statistics. The "Examples" vignette gives demonstrations of how `rsample` tools can be used.  

The package is still in development and is not yet on CRAN. To install it, use:
```{r install, eval = FALSE}
if (packageVersion("devtools") < 1.6) {
  install.packages("devtools")
}
devtools::install_github("topepo/rsample")
```

Note that resampled data sets created by `rsample` do not contain much overhead in memory. Since the original data is not modified, R does not make an automatic copy. 

For example, creating 50 bootstraps of a data set does not create an object that is 50-fold larger in memory:

```r
> library(rsample)
> library(mlbench)
> library(pryr)
> 
> data(LetterRecognition)
> 
> object_size(LetterRecognition)
2.64 MB
> 
> boots <- bootstraps(LetterRecognition, times = 50)
> 
> object_size(boots)
8.14 MB
> object_size(boots)/object_size(object_size)
6.92 kB
```

The memory usage for 50 boostrap samples is roughly 7-fold more than the original data set. 

