## Terminology

We define a _resample_ as a result of a two-way split of a data set. For example, when bootstrapping, one part of the resample is a sample with replacement of the original data. The other part are the instances that we not contained in the bootstrap sample. 

## `rset` Objects Contain Many Resamples

The main class in the package (`rset`) is for a _set_ or _collection_ of resamples. In 10-fold cross-validation, the set would consist of the 10 different resamples of the original data. 

Like `modelr`, the resamples are stored in data-frame-like `tibble` object. As a simple example, here is a small set of bootstraps of the `mtcars` data:

```{r mtcars_bt}
library(rsample)
set.seed(8584)
bootstraps <- boot(mtcars, times = 3)
bootstraps
```

We can see the way that the object stores the resamples using `splits`:

```{r mtcars_splits}
splits(bootstraps)
```

## Individual Resamples are `rsplit` Objects

The resamples are stored in the `splits` column in an object that has class `rsplit`. 

In this package we use the following terminology for the two partitions that comprise a resample: 

 * The _analysis_ data are those that we selected in the resample. For a bootstrap, this is the sample with replacement. For 10-fold cross-validation, this is the 90% of the data. These data are often used to fit a model or calculate a statistic in traditional bootstrapping. 
 * The _assessment_ data are usually the section of the original data not covered by the analysis set. Again, in 10-fold CV, this is the 10% held out. These data are often used to assess the performance of a model fit to the analysis data. 

While some might use the term "training" and "testing" for these data sets, we avoid them since they often conflicts with the data that result from an initial partition of the data (i.e. before resampling) 

Let's look at one of the `rsplit` objects
```{r rsplit}
first_resample <- splits(bootstraps, .elem = "splits")[[1]]
first_resample
```
This indicates that there were `r dim(splits(bootstraps, .elem = "splits")[[1]])["analysis"]` data points in the analysis set, `r dim(splits(bootstraps, .elem = "splits")[[1]])["assessment"]` instances were in the assessment set, and that the original data contained `r dim(splits(bootstraps, .elem = "splits")[[1]])["n"]` data points. The can also be determined using the `dim` function on an `rsplit` object. 

To obtain either of these data sets from an `rsplit`, the `as.data.frame` function can be used. By default, the analysis set is returned but the `data` option can be used to return the assessent data: 
```{r rsplit_df}
head(as.data.frame(first_resample))
as.data.frame(first_resample, data = "assessment")
```
