---
title: "Bootstrap Confidence Intervals"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Bootstrap Confidence Intervals}
output:
  knitr:::html_vignette:
    toc: yes
---


```{r setup, include=FALSE}
library(tidymodels)
library(nlstools)
library(GGally)
theme_set(theme_bw())
```

The bootstrap was originally intended for estimating confidence intervals for complex statistics whose variance properties are difficult to analytically derive. Davison and Hinkley's [_Bootstrap Methods and Their Application_](https://www.cambridge.org/core/books/bootstrap-methods-and-their-application/ED2FD043579F27952363566DC09CBD6A) is a great resource for these methods. `rsample` contains a few function to compute the most common types of intervals. 

## A nonlinear regression example

To demonstrate the computations for the different types of intervals, we'll use a nonlinear regression example from [Baty _et al_ (2015)](https://www.jstatsoft.org/article/view/v066i05). They showed data that monitored oxygen uptake in a patient with rest and exercise phases (in the data frame `O2K`). 

```{r O2K-dat}
library(tidymodels)
library(nlstools)
library(GGally)

data(O2K)

ggplot(O2K, aes(x = t, y = VO2)) + 
  geom_point()
```

The authors fit a segmented regression model where the transition point was known (this is the time when exercise commenced). Their model was:

```{r O2K-fit}
nonlin_form <-  
  as.formula(
    VO2 ~ (t <= 5.883) * VO2rest + 
      (t > 5.883) * 
      (VO2rest + (VO2peak - VO2rest) * (1 - exp(-(t - 5.883) / mu)))
    )

# Starting values from visual inspection
start_vals <- list(VO2rest = 400, VO2peak = 1600, mu = 1)

res <- nls(nonlin_form, start = start_vals, data = O2K) 

tidy(res)
```


`broom::tidy()` returns our analysis object in a standardized way. The column names shown here are used for most types of objects and this allows us to use the results more easily. 

For `rsample`, we'll rely on the `tidy()` method to work with bootstrap estimates when we need confidence intervals. There's an example at the end of a univariate statistic that isn't automatically formatted with `tidy()`. 

To run our model over different bootstraps, we'll write a function that uses the `split` object as input and produces a tidy data frame:

```{r model-info}
# Will be used to fit the models to different bootstrap data sets:
fit_fun <- function(split, ...) {
  # We could check for convergence, make new parameters, etc.
  nls(nonlin_form, data = analysis(split), ...) %>%
    tidy()
}
```

First, let's create a set of resamples and fit separate models to each. The options `apparent = TRUE` will be set. This creates a final resample that is a copy of the original (unsampled) data set. This is required for some of the interval methods. 

```{r resample}
set.seed(462)
nlin_bt <-
  bootstraps(O2K, times = 2000, apparent = TRUE) %>%
  mutate(models = map(splits, ~ fit_fun(.x, start = start_vals)))
nlin_bt

nlin_bt$models[[1]]
```

Let's look at the data and see if there any outliers or aberrant results:

```{r extract}
library(tidyr)
nls_coef <- 
  nlin_bt %>%
  dplyr::select(-splits) %>%
  # Turn it into a tibble by stacking the `models` col
  unnest() %>%
  # Get rid of unneeded columns
  dplyr::select(id, term, estimate) 

head(nls_coef)
```
Now let's create a scatterplot matrix:

```{r splom} 
nls_coef %>%
  # Put different parameters in columns
  tidyr::spread(term, estimate) %>% 
  # Keep only numeric columns
  dplyr::select(-id) %>% 
  ggscatmat(alpha = .25)
```

One potential outlier on the right for `VO2peak` but we'll leave it in. 

The univariate distributions are:

```{r hists}
nls_coef %>% 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 20, col = "white") + 
  facet_wrap(~ term, scales = "free_x")
```

### Percentile intervals

The most basic type of interval uses _percentiles_ of the resampling distribution. To get the percentile intervals, the `rset` object is passed as the first argument and the second argument is the list column of tidy results: 

```{r pctl}
p_ints <- int_pctl(nlin_bt, models)
p_ints
```

When overlaid with the univariate distributions: 

```{r pctl-plot}
nls_coef %>% 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 20, col = "white") + 
  facet_wrap(~ term, scales = "free_x") + 
  geom_vline(data = p_ints, aes(xintercept = .lower), col = "red") + 
  geom_vline(data = p_ints, aes(xintercept = .upper), col = "red")
```

How do these intervals compare to the parametric asymptotic values?

```{r int-compare}
parametric <- 
  tidy(res, conf.int = TRUE) %>% 
  dplyr::select(
    term,
    .lower = conf.low,
    .estimate = estimate,
    .upper = conf.high
  ) %>% 
  mutate(
    .alpha = 0.05,
    .method = "parametric"
  )

intervals <- 
  bind_rows(parametric, p_ints) %>% 
  arrange(term, .method)
intervals %>% split(intervals$term)
```

The percentile intervals are wider than the parametric intervals (which assume asymptotic normality). Do the estimates appear to be normally distributed? We can look at quantile-quantile plots:  

```{r qqplot}
nls_coef %>% 
  ggplot(aes(sample = estimate)) + 
  stat_qq() +
  stat_qq_line(alpha = .25) + 
  facet_wrap(~ term, scales = "free") 
```

### t-intervals

Bootstrap _t_-intervals are estimated by computing intermediate statistics that are _t_-like in structure. To use these, we require the estimated variance _for each individual resampled estimate_. In our example, this comes along with the fitted model object. We can extract the standard errors of the parameters. Luckily, most `tidy()` provide this in a column named `std.error`. 

The arguments for these intervals are the same:

```{r t-ints}
t_stats <- int_t(nlin_bt, models)
intervals <- 
  bind_rows(intervals, t_stats) %>% 
  arrange(term, .method)
intervals %>% split(intervals$term)
```


### Bias-corrected and accelerated intervals

For bias-corrected and accelerated (BCa) intervals, an additional argument is required. The `.fn` argument is a function that computes the statistic of interest. The first argument should be for the `rsplit` object and other arguments can be passed in using the ellipses. 

These intervals use an internal leave-one-out resample to compute the Jackknife statistic and will recompute the statistic for _every bootstrap resample_. If the statistic is expensive to compute, this may take some time. For those calculations, we use the `furrr` package so these can be computed in parallel if you have set up a parallel processing plan (see `?future::plan`). 

The user-facing function takes an argument for the function and the ellipses. 

```{r bca-comp}
bias_corr <- int_bca(nlin_bt, models, .fn = fit_fun, start = start_vals)
intervals <- 
  bind_rows(intervals, bias_corr) %>% 
  arrange(term, .method)
intervals %>% split(intervals$term)
```


## No existing tidy method

In this case, your function can emulate the minimum results: 

 * a character column called `term`, 
 * a numeric column called `estimate`, and, optionally, 
 * a numeric column called `std.error`. 

The last column is only needed for `int_t`. 

Suppose we just want to estimate the fold-increase in the outcome between the 90th and 10th percentiles over the course of the experiment. Our function might look like:

```{r fold-foo}
fold_incr <- function(split, ...) {
  dat <- analysis(split)
  quants <- quantile(dat$VO2, probs = c(.1, .9))
  tibble(
    term = "fold increase",
    estimate = unname(quants[2]/quants[1]),
    # We don't know the analytical formula for this 
    std.error = NA_real_
  )
}
```

Everything else works the same as before:

```{r fold-ci}
nlin_bt <-
  nlin_bt %>%
  mutate(folds = map(splits, fold_incr))

int_pctl(nlin_bt, folds)
int_bca(nlin_bt, folds, .fn = fold_incr)
```

## Intervals for linear(ish) parametric intervals

`rsample` also contains the `reg_intervals()` function that can be used for linear regression (via `lm()`), generalized linear models (`glm()`), or log-linear survival models (`survival::survreg()` or `survival::coxph()`). This function makes it easier to get intervals for these models. 

A simple example is a logistic regression using the dementia data from the `modeldata` package: 

```{r ad-data}
data(ad_data, package = "modeldata")
```

Let's fit a model with a few predictors: 

```{r ad-model}
lr_mod <- glm(Class ~ male + age + Ab_42 + tau, data = ad_data,
              family = binomial)
glance(lr_mod)
tidy(lr_mod)
```

Let's use this model with student-t intervals: 

```{r ad-t-int}
set.seed(29832)
lr_int <- 
  reg_intervals(Class ~ male + age + Ab_42 + tau, 
                data = ad_data,
                model_fn = "glm",
                family = binomial)
lr_int
```

We can also save the resamples for plotting: 

```{r ad-t-int-plot}
set.seed(29832)
lr_int <- 
  reg_intervals(Class ~ male + age + Ab_42 + tau, 
                data = ad_data,
                keep_reps = TRUE,
                model_fn = "glm",
                family = binomial)
lr_int
```

Now we can unnest the data to use in a ggplot: 

```{r ad-plot}
lr_int %>% 
  select(term, .replicates) %>% 
  unnest(cols = .replicates) %>% 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 30) + 
  facet_wrap(~ term, scales = "free_x") + 
  geom_vline(data = lr_int, aes(xintercept = .lower), col = "red") + 
  geom_vline(data = lr_int, aes(xintercept = .upper), col = "red") + 
  geom_vline(xintercept = 0, col = "green")
```
