---
title: "Bootstrap confidence intervals"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Bootstrap confidence intervals}
output:
  knitr:::html_vignette:
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval = rlang::is_installed("ggplot2") && rlang::is_installed("modeldata")
)
```

```{r package_setup, include = FALSE}
library(tidymodels)
library(nlstools)
library(GGally)
theme_set(theme_bw())
```


The bootstrap was originally intended for estimating confidence intervals for complex statistics whose variance properties are difficult to analytically derive. Davison and Hinkley's [_Bootstrap Methods and Their Application_](https://www.cambridge.org/core/books/bootstrap-methods-and-their-application/ED2FD043579F27952363566DC09CBD6A) is a great resource for these methods. rsample contains a few functions to compute the most common types of intervals. 

## A nonlinear regression example

To demonstrate the computations for the different types of intervals, we'll use a nonlinear regression example from [Baty _et al_ (2015)](https://www.jstatsoft.org/article/view/v066i05). They showed data that monitored oxygen uptake in a patient with rest and exercise phases (in the data frame `O2K`). 

```{r O2K-dat, fig.alt = "A scatterplot with time on the x-axis and oxygen uptake on the y-axis. For time < 6, the oxygen uptake remains at around 400. Between time = 6 and time = 9, it rises to about 1600, then remains around that value."}
library(tidymodels)
library(nlstools)
library(GGally)

data(O2K)

ggplot(O2K, aes(x = t, y = VO2)) + 
  geom_point()
```

The authors fit a segmented regression model where the transition point was known (this is the time when exercise commenced). Their model was:

```{r O2K-fit}
nonlin_form <-  
  as.formula(
    VO2 ~ (t <= 5.883) * VO2rest + 
      (t > 5.883) * 
      (VO2rest + (VO2peak - VO2rest) * (1 - exp(-(t - 5.883) / mu)))
    )

# Starting values from visual inspection
start_vals <- list(VO2rest = 400, VO2peak = 1600, mu = 1)

res <- nls(nonlin_form, start = start_vals, data = O2K) 

tidy(res)
```


`broom::tidy()` returns our analysis object in a standardized way. The column names shown here are used for most types of objects and this allows us to use the results more easily. 

For rsample, we'll rely on the `tidy()` method to work with bootstrap estimates when we need confidence intervals. There's an example at the end of a univariate statistic that isn't automatically formatted with `tidy()`. 

To run our model over different bootstraps, we'll write a function that uses the `split` object as input and produces a tidy data frame:

```{r model-info}
# Will be used to fit the models to different bootstrap data sets:
fit_fun <- function(split, ...) {
  # We could check for convergence, make new parameters, etc.
  nls(nonlin_form, data = analysis(split), ...) |>
    tidy()
}
```

First, let's create a set of resamples and fit separate models to each. The options `apparent = TRUE` will be set. This creates a final resample that is a copy of the original (unsampled) data set. This is required for some of the interval methods. 

```{r resample}
set.seed(462)
nlin_bt <-
  bootstraps(O2K, times = 2000, apparent = TRUE) |>
  mutate(models = map(splits, ~ fit_fun(.x, start = start_vals)))
nlin_bt

nlin_bt$models[[1]]
```

Let's look at the data and see if there any outliers or aberrant results:

```{r extract}
library(tidyr)
nls_coef <- 
  nlin_bt |>
  dplyr::select(-splits) |>
  # Turn it into a tibble by stacking the `models` col
  unnest(cols = models) |>
  # Get rid of unneeded columns
  dplyr::select(id, term, estimate) 

head(nls_coef)
```
Now let's create a scatterplot matrix:

```{r splom, fig.alt = "The 3x3 scatter plot matrix for the variables VO2rest, VO2peak, and mu. We can see one value for VO2peak that is larger than the rest."} 
nls_coef |>
  # Put different parameters in columns
  tidyr::pivot_wider(names_from = term, values_from = estimate) |> 
  # Keep only numeric columns
  dplyr::select(-id) |> 
  ggscatmat(alpha = .25)
```

One potential outlier on the right for `VO2peak` but we'll leave it in. 

The univariate distributions are:

```{r hists, fig.alt = "Three histograms for mu, VO2peak, and VO2rest: all distributions are unimodal and symmetric."}
nls_coef |> 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 20, col = "white") + 
  facet_wrap(~ term, scales = "free_x")
```

### Percentile intervals

The most basic type of interval uses _percentiles_ of the resampling distribution. To get the percentile intervals, the `rset` object is passed as the first argument and the second argument is the list column of tidy results: 

```{r pctl}
p_ints <- int_pctl(nlin_bt, models)
p_ints
```

When overlaid with the univariate distributions: 

```{r pctl-plot, fig.alt = "The three histograms for mu, VO2peak, and VO2rest with added vertical lines for the estimated lower and upper bounds of the percentile intervals."}
nls_coef |> 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 20, col = "white") + 
  facet_wrap(~ term, scales = "free_x") + 
  geom_vline(data = p_ints, aes(xintercept = .lower), col = "red") + 
  geom_vline(data = p_ints, aes(xintercept = .upper), col = "red")
```

How do these intervals compare to the parametric asymptotic values?

```{r int-compare}
parametric <- 
  tidy(res, conf.int = TRUE) |> 
  dplyr::select(
    term,
    .lower = conf.low,
    .estimate = estimate,
    .upper = conf.high
  ) |> 
  mutate(
    .alpha = 0.05,
    .method = "parametric"
  )

intervals <- 
  bind_rows(parametric, p_ints) |> 
  arrange(term, .method)
intervals |> split(intervals$term)
```

The percentile intervals are wider than the parametric intervals (which assume asymptotic normality). Do the estimates appear to be normally distributed? We can look at quantile-quantile plots:  

```{r qqplot, fig.alt = "Three QQ plots for mu, VO2peak, and VO2rest which show that the empirical distributions deviate from a normal distribution in the tails."}
nls_coef |> 
  ggplot(aes(sample = estimate)) + 
  stat_qq() +
  stat_qq_line(alpha = .25) + 
  facet_wrap(~ term, scales = "free") 
```

### t-intervals

Bootstrap _t_-intervals are estimated by computing intermediate statistics that are _t_-like in structure. To use these, we require the estimated variance _for each individual resampled estimate_. In our example, this comes along with the fitted model object. We can extract the standard errors of the parameters. Luckily, most `tidy()` methods provide this in a column named `std.error`. 

The arguments for these intervals are the same:

```{r t-ints}
t_stats <- int_t(nlin_bt, models)
intervals <- 
  bind_rows(intervals, t_stats) |> 
  arrange(term, .method)
intervals |> split(intervals$term)
```


### Bias-corrected and accelerated intervals

For bias-corrected and accelerated (BCa) intervals, an additional argument is required. The `.fn` argument is a function that computes the statistic of interest. The first argument should be for the `rsplit` object and other arguments can be passed in using the ellipses. 

These intervals use an internal leave-one-out resample to compute the Jackknife statistic and will recompute the statistic for _every bootstrap resample_. If the statistic is expensive to compute, this may take some time. For those calculations, we use the furrr package so these can be computed in parallel if you have set up a parallel processing plan (see `?future::plan`). 

The user-facing function takes an argument for the function and the ellipses. 

```{r bca-comp}
bias_corr <- int_bca(nlin_bt, models, .fn = fit_fun, start = start_vals)
intervals <- 
  bind_rows(intervals, bias_corr) |> 
  arrange(term, .method)
intervals |> split(intervals$term)
```


## No existing tidy method

In this case, your function can emulate the minimum results: 

 * a character column called `term`, 
 * a numeric column called `estimate`, and, optionally, 
 * a numeric column called `std.error`. 

The last column is only needed for `int_t()`. 

Suppose we just want to estimate the fold-increase in the outcome between the 90th and 10th percentiles over the course of the experiment. Our function might look like:

```{r fold-foo}
fold_incr <- function(split, ...) {
  dat <- analysis(split)
  quants <- quantile(dat$VO2, probs = c(.1, .9))
  tibble(
    term = "fold increase",
    estimate = unname(quants[2]/quants[1]),
    # We don't know the analytical formula for this 
    std.error = NA_real_
  )
}
```

Everything else works the same as before:

```{r fold-ci}
nlin_bt <-
  nlin_bt |>
  mutate(folds = map(splits, fold_incr))

int_pctl(nlin_bt, folds)
int_bca(nlin_bt, folds, .fn = fold_incr)
```

## Intervals for linear(ish) parametric intervals

rsample also contains the `reg_intervals()` function that can be used for linear regression (via `lm()`), generalized linear models (`glm()`), or log-linear survival models (`survival::survreg()` or `survival::coxph()`). This function makes it easier to get intervals for these models. 

A simple example is a logistic regression using the dementia data from the `modeldata` package: 

```{r ad-data}
data(ad_data, package = "modeldata")
```

Let's fit a model with a few predictors: 

```{r ad-model}
lr_mod <- glm(Class ~ male + age + Ab_42 + tau, data = ad_data,
              family = binomial)
glance(lr_mod)
tidy(lr_mod)
```

Let's use this model with student-t intervals: 

```{r ad-t-int}
set.seed(29832)
lr_int <- 
  reg_intervals(Class ~ male + age + Ab_42 + tau, 
                data = ad_data,
                model_fn = "glm",
                family = binomial)
lr_int
```

We can also save the resamples for plotting: 

```{r ad-t-int-plot}
set.seed(29832)
lr_int <- 
  reg_intervals(Class ~ male + age + Ab_42 + tau, 
                data = ad_data,
                keep_reps = TRUE,
                model_fn = "glm",
                family = binomial)
lr_int
```

Now we can unnest the data to use in a ggplot: 

```{r ad-plot, fig.alt = "Four histograms for the distribution of the regression model coefficients, with vertical lines indicating the estimated lower and upper interval bounds, as well as vertical lines for a coefficient value of 0. For age, the interval encloses 0, for the other variables it does not."}
lr_int |> 
  select(term, .replicates) |> 
  unnest(cols = .replicates) |> 
  ggplot(aes(x = estimate)) + 
  geom_histogram(bins = 30) + 
  facet_wrap(~ term, scales = "free_x") + 
  geom_vline(data = lr_int, aes(xintercept = .lower), col = "red") + 
  geom_vline(data = lr_int, aes(xintercept = .upper), col = "red") + 
  geom_vline(xintercept = 0, col = "green")
```
